---
title: "Skip negative control adjustment"
author: "Candace Savonen"
date: "2025-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up data

We're going to set up three datasets that we will provide to the `set_up()` function to create a `gimap` dataset object.

- `counts` - the counts generated from pgPEN
- `pg_ids` - the IDs that correspond to the rows of the counts and specify the construct
- `sample_metadata` - metadata that describes the columns of the counts including their timepoints

```{r}
example_data <- get_example_data("count")

counts <- example_data %>%
  select(c("Day00_RepA", "Day05_RepA", "Day22_RepA", "Day22_RepB", "Day22_RepC")) %>%
  as.matrix()

pg_ids <- example_data %>%
  dplyr::select("id")
```

Remove NTs for one of the datasets.

```{r}
pg_ids_nt <- grep("NTpg", pg_ids$id, invert = TRUE)

counts_nt <- counts[pg_ids_nt, ]
pg_ids_nt <- pg_ids[pg_ids_nt, ]
```

Sample metadata is the information that describes the samples and is sorted the same order as the columns in the count data.

```{r }
sample_metadata <- data.frame(
  col_names = c("Day00_RepA", "Day05_RepA", "Day22_RepA", "Day22_RepB", "Day22_RepC"),
  day = as.numeric(c("0", "5", "22", "22", "22")),
  rep = as.factor(c("RepA", "RepA", "RepA", "RepB", "RepC"))
)
```


```{r}
gimap_dataset <- setup_data(
  counts = counts,
  pg_ids = pg_ids,
  sample_metadata = sample_metadata
)

gimap_dataset_no_nts <- setup_data(
  counts = counts_nt,
  pg_ids = pg_ids_nt,
  sample_metadata = sample_metadata
)
```

## Set up comparison sets

```{r}
 gimap_dataset <- gimap_dataset %>%
   gimap_filter() %>%
   gimap_annotate(cell_line = "HELA") %>%
   gimap_normalize(
     timepoints = "day",
     missing_ids_file =  tempfile()
   ) %>%
   calc_gi()

 gimap_dataset_ratio <- gimap_dataset_no_nts %>%
   gimap_filter() %>%
   gimap_annotate(cell_line = "HELA") %>%
   gimap_normalize(
     timepoints = "day",
     missing_ids_file =  tempfile(),
     no_neg_ctrl_adj = TRUE
   ) %>%
   calc_gi()

 
gimap_dataset_no_adj <- gimap_dataset_no_nts %>%
  gimap_filter() %>%
  gimap_annotate(cell_line = "HELA") %>%
  gimap_normalize(
    timepoints = "day",
    missing_ids_file =  tempfile(),
     no_adj = TRUE
  ) %>%
  calc_gi()
```
 
 
```{r}
 plot_exp_v_obs_scatter(gimap_dataset, 
                        facet_rep = FALSE, 
                        reps_to_drop = "Day05_RepA_early")

 plot_exp_v_obs_scatter(gimap_dataset_no_nts,
                        facet_rep = FALSE,
                        reps_to_drop = "Day05_RepA_early")
```
 
 
```{r}
 plot_rank_scatter(gimap_dataset, 
                   reps_to_drop = "Day05_RepA_early")
 
 plot_rank_scatter(gimap_dataset_no_nts, 
                   reps_to_drop = "Day05_RepA_early")
```
 
```{r}
 plot_volcano(gimap_dataset, reps_to_drop = "Day05_RepA_early")
 plot_volcano(gimap_dataset_no_nts, reps_to_drop = "Day05_RepA_early")
```

