---
title: "scratch gimap CRISPR Score Review"
output: html_document
date: "`r Sys.Date()`"
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(devtools)
library(magrittr)
library(ggpubr)

devtools::load_all()
```

## Get the gimap CRISPR Score results

```{r}
 gimap_dataset <- get_example_data("gimap") %>%
   gimap_filter() %>%
   gimap_annotate(cell_line = "HELA") %>%
   gimap_normalize(
     timepoints = "day") %>%
  calc_crispr()
```

## Load the GI Mapping CRISPR Score results

```{r}
old_crispr_results <- read_tsv("HeLa_lfc_annot_adj_pgRNA.txt")
```
### Negative and positive control medians 

We are expecting negative controls to be now set to a median of 0. 

For the original data: 
```{r}
old_neg_controls <- old_lfc_results %>%
  dplyr::filter(target_type == "ctrl_ctrl") %>%
  dplyr::group_by(rep) %>%
  dplyr::summarize(neg_ctrl_med = median(lfc_adj2)) %>%
  dplyr::pull(neg_ctrl_med)

old_neg_controls
```

For the new data: 
```{r}
new_neg_controls <- gimap_dataset$single_crispr_score %>%
  dplyr::filter(target_type == "ctrl_ctrl", rep != "Day05_RepA_early") %>%
  dplyr::group_by(rep) %>%
  dplyr::summarize(neg_ctrl_med = median(lfc_adj)) %>%
  dplyr::pull(neg_ctrl_med)

new_neg_controls
```

We are expecting negative controls to be now set to a median of -1. 

For the original data: 
```{r}
old_pos_controls <- old_lfc_results %>%
  dplyr::filter(target_type == "gene_ctrl" & gene1_essential_flag == TRUE | 
                  target_type =="ctrl_gene" & gene2_essential_flag == TRUE) %>%
  dplyr::group_by(rep) %>%
  dplyr::summarize(pos_ctrl_med = median(CRISPR_score)) %>%
  dplyr::pull(pos_ctrl_med)

old_pos_controls
```

For the new data: 
```{r}
new_pos_controls <- gimap_dataset$single_crispr_score %>%
  dplyr::filter(rep != "Day05_RepA_early") %>%
  dplyr::group_by(rep) %>%
  dplyr::summarize(pos_ctrl_med = median(single_crispr)) %>%
  dplyr::pull(pos_ctrl_med)

new_pos_controls
```

#### sync the reps values

```{r}
old_crispr_results %<>% mutate(
  rep = recode(rep, "A" = "Day22_RepA_late", "B" = "Day22_RepB_late", "C" = "Day22_RepC_late")
)
```

## Inspect/Compare

```{r}
ncol(gimap_dataset$crispr_score)
ncol(old_crispr_results)

print("Column names in new results not in old results")
setdiff(colnames(gimap_dataset$crispr_score), colnames(old_crispr_results))
print("Column names in old results not in new results")
setdiff(colnames(old_crispr_results), colnames(gimap_dataset$crispr_score))


nrow(gimap_dataset$crispr_score)
nrow(old_crispr_results)
```

From [this line in the code](https://github.com/FredHutch/gimap/blob/273d9cb50d80b0d8c29375582eec9d1946bd7556/R/05-crispr-calc.R#L133), it appears that the `double_crispr_score` column from gimap results is related to the `CRISPR_score` column from GI_Mapping results.

Why are there so many more rows in gimap results? Some data does get dropped after the join (225993 rows before the drop_na)

There appears to be 4 entries for each pgRNA ID and rep combo. They share the same original CRISPR score but once you look at others, that seems to be where the differences are.  

```{r}
joindf <- dplyr::full_join(old_crispr_results, gimap_dataset$crispr_score,
                           by = c("pgRNA_id" = "pg_ids", "rep"),
                           suffix = c("_old", "_new")) %>%
  select(double_crispr_score, CRISPR_score, rep, pgRNA_id) %>% #225993 rows prior to drop_na
  drop_na() %>%
  distinct()

nrow(joindf)
```

```{r}
joindf %>% ggplot(aes(x=double_crispr_score, y = CRISPR_score)) +
  geom_point(alpha=0.1) +
  xlab("gimap Normalization/LFC: lfc_adj") +
  ylab("GI Mapping Normalization/LFC: lfc_adj2") +
  theme_classic() +
  stat_regline_equation(label.y = 2.35) + stat_cor(label.y = 1.95) +
  facet_wrap(~rep)
```
```{r}
joindf %>% ggplot(aes(x=double_crispr_score, y = CRISPR_score)) +
  geom_point(alpha=0.1) +
  xlab("gimap Normalization/LFC: lfc_adj") +
  ylab("GI Mapping Normalization/LFC: lfc_adj2") +
  theme_classic() +
  stat_regline_equation(label.y = 2.35) + stat_cor(label.y = 1.95)
```
