---
title: "scratch gimap Genetic Interaction Review"
output: html_document
date: "`r Sys.Date()`"
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(devtools)
library(magrittr)

devtools::load_all()
```

## Get the gimap GI results

```{r}
 gimap_dataset <- get_example_data("gimap") %>%
   gimap_filter() %>%
   gimap_annotate(cell_line = "HELA") %>%
   gimap_normalize(
     timepoints = "day") %>%
  calc_crispr() %>%
  calc_gi()
```

```{r}
saveRDS(gimap_dataset, "gimap_dataset_calculated.RDS")
```

## Load the GI Mapping GI results

```{r}
old_gi_results <- readRDS("d.HeLa_GI_scores_target")
```

#### sync the reps values

```{r}
old_gi_results %<>% mutate(
  rep = recode(rep, "A" = "Day22_RepA_late", "B" = "Day22_RepB_late", "C" = "Day22_RepC_late")
)
```

## Compare results

### Columns

```{r}
colnames(gimap_dataset$gi_scores)
```

```{r}
colnames(old_gi_results)
```

Looks like we want to compare the pgRNA targets which should be the `pgRNA_target_double` column for the gimap results and the `pgRNA_target` column for the GI Mapping results

Columns in the GI Mapping results not in the gimap results seem to include model results (`intercept`, `slope`) as well as some significance values (`p_val`, `fdr`), observed and expected CRISPR score (`mean_observed_CS` and `mean_expected_CS` (which I think I need for plotting)). And how does `mean_GI_score` relate to the three GI score columns in the gimap results? The three GI score columns relate to the different target types.  

### Target overlap

```{r}

length(unique(gimap_dataset$normalized_log_fc$pgRNA_target))

length(unique(old_gi_results$pgRNA_target))
length(unique(gimap_dataset$gi_scores$pgRNA_target_double))
length(intersect(unique(old_gi_results$pgRNA_target), unique(gimap_dataset$gi_scores$pgRNA_target_double)))
```

Only 1/3 of the GI Mapping targets are represented in the gimap results...

### Number of observations/rows total and rep column overlap

```{r}
nrow(old_gi_results)
nrow(gimap_dataset$gi_scores)
```

The gimap results has far more rows...

```{r}
table(old_gi_results$rep)

table(gimap_dataset$gi_scores$rep)
```

While the replicate values themselves can't be completely synced like we've done in other reviews due to the gimap results including the Day05_RepA_early datapoints, this further shows the far greater number of observations (even if there are fewer represented pgRNA targets) in the gimap results....

```{r}
head(old_gi_results)

```

```{r}
joindf <- dplyr::full_join(old_gi_results, gimap_dataset$gi_scores,
                           by = c("pgRNA_target" = "pgRNA_target_double", "rep"),
                           suffix = c("_old", "_new"))
```

## Split out the comparisons we want to make

Have to drop NA's because the gimap results include Day05_RepA_early datapoints while the GI Mapping doesn't. Also have to drop NAs because there are GI Mapping targets that aren't represented in the gimap targets?

```{r}
joined_df <- rbind(
  #join the gene_ctrl
  full_join(
    old_gi_results %>% filter(target_type == "gene_ctrl") %>% select(c("pgRNA_target", "paralog_pair", "mean_GI_score", "rep")),
    gimap_dataset$gi_scores %>% separate(pgRNA_target_double, c("gene1", "gene2"), sep="_", remove = FALSE) %>% mutate(ctrl = "ctrl") %>% unite("pgRNA_target_summary", c("gene1", "ctrl"), sep="_") %>% select(c("pgRNA_target_summary", "pgRNA_target_double", "single_target_gi_score_1", "rep")),
    by = c("pgRNA_target" = "pgRNA_target_summary", "paralog_pair"="pgRNA_target_double", "rep"),
    suffix = c("_old", "_new")
    ) %>% mutate(target_type = "gene_ctrl") %>% distinct() %>% select(c("pgRNA_target", "mean_GI_score", "rep", "single_target_gi_score_1", "target_type")) %>% `colnames<-`(c("pgRNA_target", "GI_Mapping_GI_score", "rep", "gimap_GI_score", "target_type")) %>% drop_na(),
  # join the ctrl_gene
  full_join(
    old_gi_results %>% filter(target_type == "ctrl_gene") %>% select(c("pgRNA_target", "paralog_pair", "mean_GI_score", "rep")),
    gimap_dataset$gi_scores %>% separate(pgRNA_target_double, c("gene1", "gene2"), sep="_", remove=FALSE) %>% mutate(ctrl = "ctrl") %>% unite("pgRNA_target_summary", c("ctrl", "gene2"), sep="_") %>% select(c("pgRNA_target_summary", "pgRNA_target_double", "single_target_gi_score_2", "rep")),
    by = c("pgRNA_target" = "pgRNA_target_summary", "rep"),
    suffix = c("_old", "_new")
    ) %>% mutate(target_type = "ctrl_gene") %>% distinct() %>% select(c("pgRNA_target", "mean_GI_score", "rep", "single_target_gi_score_2", "target_type")) %>% `colnames<-`(c("pgRNA_target", "GI_Mapping_GI_score", "rep", "gimap_GI_score", "target_type")) %>% drop_na(),
  #join the gene_gene
  full_join(
    old_gi_results %>% filter(target_type == "gene_gene") %>% select(c("pgRNA_target", "mean_GI_score", "rep")),
    gimap_dataset$gi_scores %>% select(c("pgRNA_target_double", "double_target_gi_score", "rep")),
    by = c("pgRNA_target" = "pgRNA_target_double", "rep"),
    suffix = c("_old", "_new")
  ) %>% mutate(target_type = "gene_gene") %>% `colnames<-`(c("pgRNA_target", "GI_Mapping_GI_score", "rep", "gimap_GI_score", "target_type")) %>% drop_na()
)
```

```{r}
joined_df %>% ggplot(aes(x=gimap_GI_score, y=GI_Mapping_GI_score, color=target_type)) + geom_point() + facet_wrap(rep~target_type) + theme(legend.position = "none")
```