---
title: "scratch gimap Genetic Interaction Review"
output: html_document
date: "`r Sys.Date()`"
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(devtools)
library(magrittr)

devtools::load_all()
```

## Get the gimap GI results

```{r}
gimap_dataset <- get_example_data("gimap") %>%
  gimap_filter() %>%
  gimap_annotate(cell_line = "HELA") %>%
  gimap_normalize(
    timepoints = "day"
  ) %>%
  calc_crispr() %>%
  calc_gi()
```

```{r}
saveRDS(gimap_dataset, "gimap_dataset_calculated.RDS")
```

## Load the GI Mapping GI results

```{r}
old_gi_results <- readRDS("d.HeLa_GI_scores_target")
```

#### sync the reps values

```{r}
old_gi_results %<>% mutate(
  rep = recode(rep, "A" = "Day22_RepA_late", "B" = "Day22_RepB_late", "C" = "Day22_RepC_late")
)
```

## Compare results

### Columns

```{r}
colnames(gimap_dataset$gi_scores)
```

```{r}
colnames(old_gi_results)
```

Looks like we want to compare the pgRNA targets which should be the `pgRNA_target_double` column for the gimap results and the `pgRNA_target` column for the GI Mapping results

Columns in the GI Mapping results not in the gimap results seem to include model results (`intercept`, `slope`) as well as some significance values (`p_val`, `fdr`), observed and expected CRISPR score (`mean_observed_CS` and `mean_expected_CS` (which I think I need for plotting)). And how does `mean_GI_score` relate to the three GI score columns in the gimap results? The three GI score columns relate to the different target types.  

### Target overlap

```{r}
length(unique(gimap_dataset$normalized_log_fc$pgRNA_target))
length(unique(old_gi_results$pgRNA_target))
length(unique(gimap_dataset$gi_scores$pgRNA_target))
length(intersect(unique(old_gi_results$pgRNA_target), unique(gimap_dataset$gi_scores$pgRNA_target)))
```

### Number of observations/rows total and rep column overlap

```{r}
nrow(old_gi_results)
nrow(gimap_dataset$gi_scores)
```

The gimap results has far more rows...

```{r}
table(old_gi_results$rep)

table(gimap_dataset$gi_scores$rep)
```

While the replicate values themselves can't be completely synced like we've done in other reviews due to the gimap results including the Day05_RepA_early datapoints, this further shows the far greater number of observations (even if there are fewer represented pgRNA targets) in the gimap results....

```{r}
head(old_gi_results)
```

## Split out the comparisons we want to make

Have to drop NA's because the gimap results include `Day05_RepA_early` datapoints while the GI Mapping doesn't. Also have to drop NAs because there are GI Mapping targets that aren't represented in the gimap targets?

```{r}
View(gimap_dataset$gi_scores)

# Reshape the data so we can plot
joined_df <- gimap_dataset$gi_scores %>% 
  dplyr::inner_join(old_gi_results, 
                   by = c("pgRNA_target", "rep", "target_type"), 
                   suffix = c("_new", "_old")) %>% 
  dplyr::filter(!grepl("Day05", rep))
```



```{r}
joined_df %>% 
  ggplot(aes(x = mean_gi_score, y = mean_GI_score)) +
  geom_point() +
  facet_wrap(~target_type) +
  theme(legend.position = "none")
```
